Scalar Feedback based Joint Time-Frequency Precoder Interpolation for MIMO-OFDM Systems

### 2020-03-04 paper/response_letter.tex

``\emph{Independent time-frequency based approaches correspond to the
  method without joint-time frequency interpolation i.e. with
  prediction in time domain along with separate interpolation in the frequency
  domain.}''

\textbf{2. The quantization of $\phi_{k,l}$'s and $\theta_{k,l}$'s mentioned in Subsection III.A and the interpolation
process mentioned in Subsection III.C should be explained more clearly.}

\textbf{Response:}
Thanks for pointing this out. The explanation of the quantization
process has now been made more  clear by altering the text in III.A as
follows:

``\emph{If $n$ bits are allocated to quantize a parameter
for the first time, the range of values that each parameter takes is
divided into $2^n$ optimal intervals using Lloyd's
quantization [17].}''

@@

\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future
  precoder parameters are required to conduct the interpolation. How
  can we obtain the future precoder parameters?}

\textbf{Response:} We thank the reviewer for pointing this out this
missing piece of information. We
wish to clarify that the precoder used for transmission is used for
transmission. However, the past precoder estimates are improved using
the new information, and this provides slight improvement when
performing prediction in the further steps. This was not stated
clearly in the previous manuscript. We have now altered the text in the subsection III.C as follows:

\emph{To construct the rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' and time
instances to do the interpolation (using backtracking to ensure
causality). By future time instances it means that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy.}

@@

In addition, the temporal correlation based precoder tracking
discussed in [8] and [9] have considered about precoder tracking in
time and frequency respectively, albeit separately, without any joint
interpolation.  We have approached this differently by performing
prediction in the time domain and using this for joint interpolation
on the time-freqeuncy grid. This helps build a framework for reducing
the total bit bandwidth by a significant amount while maintaining
similar performance in terms of BER.  We have altered the text in the
introduction to make this point more clear, as follows:

``\emph{In our work, we perform prediction in time domain, and use this for joint
  interpolation on the time frequency grid.}''

This is in addition to various changes in other sections of the paper
that further emphasize the advantages of our approach.

\textbf{6. The reference format is not consistent and there are
  several typos. The title of the manuscript is not in IEEE format.}

Thank you for the comment. We have now improved the references and
fixed the errors.
\end{document}

### 2020-03-13 paper/response_letter.tex

\textbf{1. Work/works got mixed up for many times.}

Thank you for this comment. We have fixed all instances of this error.

\textbf{2. ``The quantization approach described above... '' Actually,
  parameterization is described without presenting the quantizing
  method.}

\textbf{Response:} We thank the reviewer for pointing out this
error. We have reworded Section III to clarify the parameterization and quantization
in a suitable order. In addition, we have altered the text to include this:

@@

\textbf{2. The quantization of $\phi_{k,l}$'s and $\theta_{k,l}$'s mentioned in Subsection III.A and the interpolation
process mentioned in Subsection III.C should be explained more clearly.}

\textbf{Response:}
Thanks for pointing this out. The explanation of the quantization
process has now been made more  clear by altering the text in III.A as
follows:

``\emph{If $n$ bits are allocated to quantize a parameter
for the first time, the range of values that each parameter takes is
divided into $2^n$ optimal intervals using Lloyd's
quantization [17].}''

In addition we have clarified the interpolation process by precisely stating
how we are combining temporaly separated precoders as follows:

\emph{To construct precoders for rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' time
instances to do the interpolation (using backtracking to ensure
causality). When using future time instances, the implication is that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy.}

### 2020-03-13 paper/main.tex
make changes

\title{Scalar Feedback based Joint Time-Frequency Precoder
  interpolation for  MIMO-OFDM Systems}

@@

An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders. A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{mondal2007quantization,schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}. On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}. While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
involve operations over higher dimensional manifolds. The scalar parametrization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
Such methods have also been implemented in 802.11 ac and ax standards
\cite{lou2013comparison,ieee80211}. We are not aware of performance comparisons
between these two bodies of work, thereby making it difficult to
quantify from past work whether performance gains/losses can accrue by operating over scalar parametrization instead of manifolds. In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques. One of the key results of our work is that the proposed scalar parametrization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.

@@

represented in Fig.~\ref{fig:adpm-fig}. It is known~\cite{4114278}

@@

For the adaptive channel tracking scheme, the arrows in
Fig.~\ref{fig:adpm-fig} show how previous quantized values along with
the 1-bit enhancement are used to find the new quantized value. Since
we use a single bit adaptation, the amount of feedback required is 1
bit for each scalar parameter that is being tracked. On average, this
can be reduced by half if we transmit only $\bf{\Theta}$ values for
one-time instance and only $\bf{\Phi}$ values for the next time
instant in an alternating fashion (the untransmitted value is assumed
to be the same as that at the previous instant). In this case, the
average number of bits required to quantize a channel matrix will be
$N_{R}(2N_{T} - N_R )/2$ (from~\cite{4114278}). While tracking
the $\phi_{k,l}$ parameters, it must be noted they may change abruptly
between $-\pi$ and $\pi$ due to jumps of $2\pi$. This is avoided by
unwrapping the phases to facilitate continuous tracking (i.e., jumps
in phase of magnitude close to $2\pi$ have been eliminated).
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$, in keeping with the same notation as the previous
section. To construct the rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' and time
instances to do the interpolation (using backtracking to ensure
causality). By future time instances it means that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
using the neighbouring subcarriers using bilinear interpolation
(equivalent to linear interpolation in two dimensions). These
interpolated values are then used to reconstruct the precoding
matrices all together by applying the reverse transformation of
breaking the precoder into independent parameters as given in
Equation~\ref{eq:decompose}. Thus, the approach here is able to
exploit the time and frequency correlation jointly to enhance
performance.

>>>>

\title{Scalar Feedback based Joint Time-Frequency Precoder
  Interpolation for  MIMO-OFDM Systems}

@@

An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders. A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}. On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}. While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
involve operations over higher dimensional manifolds. The scalar parametrization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
Such methods have also been implemented in 802.11 ac and ax standards
\cite{lou2013comparison,ieee80211}. We are not aware of performance comparisons
between these two bodies of work, thereby making it difficult to
quantify from past work whether performance gains/losses can accrue by operating over scalar parametrization instead of manifolds. In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques. 
One of the key results of our work is that the proposed scalar parametrization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
With smaller quantization error as compared to manifold based approaches, the feedback
results in more accurate precoder tracking, along with simple 2D-interpolation over the
scalar parameters to optimize the precoders

@@

represented in Fig.~\ref{fig:adpm-fig}. It is known 

@@

For the adaptive channel tracking scheme, the arrows in
Fig.~\ref{fig:adpm-fig} show how previous quantized values along with
the 1-bit enhancement are used to find the new quantized value. Since
we use a single bit adaptation, the amount of feedback required is 1
bit for each scalar parameter that is being tracked. On average, this
can be reduced by half if we transmit only $\bf{\Theta}$ values for
one-time instance and only $\bf{\Phi}$ values for the next time
instant in an alternating fashion (the untransmitted value is assumed
to be the same as that at the previous instant). In this case, the
average number of bits required to quantize a channel matrix will be
$N_{R}(2N_{T} - N_R )/2$ (from~\cite{4114278}). While tracking
the $\phi_{k,l}$ parameters, it must be noted they may change abruptly
between $-\pi$ and $\pi$ due to jumps of $2\pi$. This is avoided by
unwrapping the phases to facilitate continuous tracking (i.e., jumps
in phase of magnitude close to $2\pi$ have been eliminated).
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$, keeping with the same notation as the previous
section. To construct precoders for rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' time
instances to do the interpolation (using backtracking to ensure
causality). When using future time instances, the implication is that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
using the neighbouring subcarriers using bilinear interpolation
(equivalent to linear interpolation in two dimensions). These
interpolated values are then used to reconstruct the precoding
matrices all together by applying the reverse transformation of
breaking the precoder into independent parameters as given in
Equation~\ref{eq:decompose}. Thus, the approach here is able to
exploit the time and frequency correlation jointly to enhance
performance.
